---
title: '**Data quality report for REDCap data**'
author: "By Kirstine K. Rasmussen and Sophia Hejndorf"
date: "Created `r format(Sys.time(), '%d-%m-%Y %X')`"
output:
  html_document:
    css: assets/MISTRAL_CSS_style.css
    toc: yes
    toc_float: yes
    toc_depth: 2
    collapsed: no
    theme: simplex
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "output") })
---

<script>
  $(document).ready(function() {
    $('#TOC').parent().prepend('<div id=\"nav_logo\"><img src=\"assets/MISTRAL_logo.jpg\"></div>');
  });
</script>

<style>
#TOC {
  margin-top: 120px;
}
#nav_logo {
  position: fixed;
  width: 15%;
  margin-top: 10px;
}
</style>

<br>

*This data quality report has been created for publishing. It is a shorter, simplified version of the report used for quality assurance (QA) in the Horizon 2020 funded observational study [MISTRAL](https://www.chip.dk/Research/Studies/MISTRAL). All data has been annonymised and randomised, and only selected sections are included to showcase the usage of the report.*

<br>

```{r setup, include=FALSE}

# Define packages to load (packages needs to be installed using install.packages() before they can be loaded)
packages <- c("plotly","plyr","tidyverse","DT","data.table","janitor","knitr","scales","Hmisc","httr",
              "usethis","odbc","DBI","stringr","dplyr","lubridate","heatmaply","here")

# Check if packages are installed
missing_packages <- packages[sapply(packages, function(p) system.file(package = p) == "")]

# Install packages not installed
if(length(missing_packages) > 0) {
  install.packages(missing_packages, dependencies = TRUE)
}

# Load all packages
lapply(packages, require, character.only = TRUE)

# Set up document
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 9) # R Markdown setup
options(scipen = 999) # Disable scientific notation (e+09)
options(dplyr.summarise.inform = FALSE) # Disable messages for dplyr

```

```{r style_setup, include=FALSE}

# Define colours to use in all of the report
colPositive <-  "#6d1c68"
colNegative <-  "#2d6987"
colUnknown <-   "#fde333"
colNA <-        "#969696"
colYes <-       "#30893b"
colNo <-        "#cc0000"
colMISTRAL <-   "#E64659"
colCD4 <-       "#e28743"
colCD8 <-       "#0e79af"

# Define styling for tables
table_style <- 'cell-border stripe hover compact'

```

```{r Get data through API, include=FALSE}
# #Set up function for API connection to REDCap
# #NB: Not actively used in the dummy report, but code made available for the reader to implement in their own study.
# API_get_data <- function(token) {
#   # Define URL for data retrieval
#   url <- "https://www.my-url.com"
#   # Set up certificate verification (safety step)
#   httr::set_config(httr::config(verbose = FALSE, ssl_verifypeer = TRUE, ssl_verifyhost = 2))
#   # Get data using specific url and token
#   result <- redcap_read(redcap_uri = url, token = token)$data
#   # Return downloaded data
#   return(result)
# }
```


```{r Get data from local data file, include=FALSE}

# Load R-formatted Visits form data
source("src/formatting_scripts/DataFormatting_Visits.R")

# Save data from each visit in separate dataframes
Visit1 <- Visits[Visits$redcap_event_name=="visit_1_arm_1",]
Visit2 <- Visits[Visits$redcap_event_name=="visit_2_arm_1",]

# Save visit dataframes in Rdata objects for use in child documents
save(Visit1, file="data/derived/Visit1.Rdata")
save(Visit2, file="data/derived/Visit2.Rdata")

# Load R-formatted Clinical form data
source("src/formatting_scripts/DataFormatting_Clinical.R")

```


```{r Child documents, child=c('src/child_scripts/Child_VisitsForm.Rmd', 'src/child_scripts/Child_ClinicalForm.Rmd')}
# This chunk should not contain code
# Used to call child documents with sections for each REDCap form
#
```
