---
title: "Child_Visits"
author: "Kirstine K. Rasmussen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## **Visits form** {.tabset}

*The Visits form was used for collecting information related to the participants' visit at their designated clinical site (centre). At each visit the clinical staff fills out a questionnaire as well as the sample IDs for the samples collected at that visit. Additionally, at their first visit (visit 1), a questionnaire about participant baseline data was collected.*

**Total participants who have done Visit 1: **`r nrow(Visit1)` <br> 

**Total participants who have done Visit 2: **`r nrow(Visit2)` <br><br> 

### General overview {.tabset}

#### Visit 1

**Completion date for section *Participant baseline* data** <br><br>
Min:
`r min(Visit1$es_completed_date_a1, na.rm = TRUE)` <br> 
Max:
`r max(Visit1$es_completed_date_a1, na.rm = TRUE)`

```{r, echo=FALSE}

# Plot participants enrolled at each clinical site
visitp <- ggplot(Visit1, aes(x = centre, fill = centre_name)) + 
  geom_bar() +
  theme_light() +
  xlab("Centre code") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_y_continuous(breaks = seq(0,700,10)) +
  scale_fill_discrete(name = "Centre name") +
  ggtitle("Enrolments per centre");ggplotly(visitp, height=350)

```


```{r, echo=FALSE}

# Create cummulative plot of enrolments per month

# Enrolment status
Enrolment_status <- Visit1[c("bas_patient","centre","centre_name","centre_count","bas_enrol_mistral_d")] %>%
  # Enrolment date column
  mutate(enrolment_date = as.Date(bas_enrol_mistral_d, "%Y-%m-%d")) %>%
  # Enrolment month column
  mutate(Month = as.Date(cut(enrolment_date, breaks="month"))) %>%
  # Sort by month
  arrange(Month) %>%
  # Add cummulative sums column
  mutate(quant = 1) %>%
  mutate(cum_sum = cumsum(quant))

cummulative <- setNames(aggregate(Enrolment_status$cum_sum, by = list(Enrolment_status$Month), max),
                        c("Month","cum"))

p_enrol <- ggplot(Enrolment_status, aes(x=Month)) + 
  geom_bar(stat = "count", fill=colMISTRAL) +
  scale_x_date(labels=label_date("%Y %m"), breaks="1 month") +
  xlab("Enrolments per month") +
  ylab("Participants enrolled") +
  theme_light()

# Second y-axis for cummulative
ay <- list(
  tickfont = list(size=11),
  titlefont=list(size=13),
  overlaying = "y",
  nticks = 5,
  side = "right",
  title = "Second y axis"
)

p_cum <- ggplot(Enrolment_status, aes(x=Month)) + 
  theme_light() +
  geom_bar(stat = "count", fill=colMISTRAL) +
  geom_line(data=cummulative, aes(x=Month,y=cum/10), linewidth=0.50, color="black") +
  geom_hline(yintercept = 100, linewidth=0.75, color="grey40") +
  
  scale_x_date(labels=label_date("%Y %m"), breaks="1 month") +
  xlab("Enrolments per month") +
  ggtitle("Cummulative enrolment") +
  
  scale_y_continuous(
    limits = c(0,100.7),
    breaks=seq(0,100,25),
    # Features of the first axis
    name = "Participants enrolled",
    # Add a second axis and specify its features
    sec.axis = sec_axis( transform=~.*10, name="Cummulative enrolment") ) +
  
  geom_text(stat= "count", aes(label=after_stat(count)), size = 4, nudge_y = 3) +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = 0),
        axis.title.y = element_text(color = colMISTRAL, size=13),
        axis.title.y.right = element_text(color = "grey40", size=13),
        axis.text = element_text(size=11)) +
  theme()

p_cum

```

```{r, echo=FALSE}

# Plot enrolments per month
ggplotly(p_enrol + ggtitle("Number enrolled: per centre per month") +
           facet_wrap("centre", ncol=3) +
           geom_line(y=50) +
           ylim(0,30) +
           theme(axis.text.x = element_text(angle = -45, vjust = 1, hjust=1, size = 6)), height = 800)

# Clean up unnecessary variables
#rm(Enrolment_status, p_cum, p_enrol)

```

**Last enrolment per centre**<br><br>
Table used to check when the clinical sites (centres) were last active updating REDCap.<br>

```{r, echo=FALSE}

# Display last enrolment activity per clinical site
setNames(aggregate(Visit1$es_completed_date_a1, by = list(Visit1$centre,Visit1$centre_name), max), c("Centre","Centre name","Last enrolment")) %>%
  datatable(.,rownames=FALSE, options = list(dom = 't', columnDefs = list(list(className = 'dt-center', targets = "_all")), pageLength = 25), class = table_style)

```

**Enrolments per country**<br><br>
Multiple centres were placed in the same country, so this graphic provided different insights from what was available in the REDCap forms. The country variable was added in the data processing R script.<br>

```{r, echo=FALSE}

# Plot enrolments per country
ggplotly(ggplot(Visit1, aes(x = centre_country, fill = centre_country)) + 
  geom_bar() +
  theme_light() +
  xlab("") +
  ylab("Count") + 
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_y_continuous(breaks = seq(0,1500,25)) +
  scale_fill_discrete(name = "Centre country") +
  ggtitle("Enrolments per country"), height=350)

```

```{r, echo=FALSE}

# Create dataframe of promised enrolments per clinical site
site_promised_numbers <- Visit1[c("centre","centre_count","centre_name")] %>% unique(.)

```

**Total promised enrolled participants from centres who have started enrolling:** `r sum(site_promised_numbers$centre_count)`<br><br>
Showing the number of enrolled participants relative to the number they promised they could enrol. <br>

```{r, echo=FALSE}

# Plot enrolments per centre and include the number of promised enrolments
ggplotly(ggplot(Visit1, aes(x = centre, fill = centre_name)) + 
  geom_bar() +
  geom_point(data = site_promised_numbers, mapping=aes(x=centre,y=centre_count), stroke=0.2) +
  theme_light() +
  xlab("Centre code") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_y_continuous(breaks = seq(0,700,10)) +
  scale_fill_discrete(name = "Centre name") +
  ggtitle("Enrolments per centre and their promised number of enrolled"), height=350)

```

<br><br>

#### Visit 2 

**Completion date for section *Questionnaire* data** <br><br>
Min:
`r min(Visit2$es_completed_date_a1b, na.rm = TRUE)` <br> 
Max:
`r max(Visit2$es_completed_date_a1b, na.rm = TRUE)`

```{r, echo=FALSE}

# Plot number of participants who completed visit 2 at each clinical site
visitp <- ggplot(Visit2, aes(x = centre, fill = centre_name)) + 
  geom_bar() +
  theme_light() +
  xlab("Centre code") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_y_continuous(breaks = seq(0,700,10)) +
  scale_fill_discrete(name = "Centre name") +
  ggtitle("Follow-up at visit 2 per centre");ggplotly(visitp, height=350)

```

**Last visit 2 per centre**<br><br>
Table used to check when the clinical sites (centres) were last active updating REDCap.<br>
Empty cells represent clinical sites who have not yet started their visit 2s.<br>

```{r, echo=FALSE}

# Display when last visit 2 happened per clinical site
setNames(aggregate(Visit2$es_completed_date_a1b, by = list(Visit2$centre,Visit2$centre_name), max), 
         c("Centre","Centre name","Last visit 2 date")) %>%
  datatable(.,rownames=FALSE, class = table_style, 
            options = list(dom = 't', columnDefs = list(list(className = 'dt-center', targets = "_all")), pageLength = 20))
```

**Participats with visit 2 per country**<br><br>

```{r, echo=FALSE}

# Plot number of participants who have had visit 2 per clinical site
ggplotly(ggplot(Visit2, aes(x = centre_country, fill = centre_country)) + 
  geom_bar() +
  theme_light() +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_y_continuous(breaks = seq(0,1500,10)) +
  scale_fill_discrete(name = "Centre country") +
  ggtitle("Visit 2 per country"), height=350)

```

<br><br><br>

### Follow up

**Dates for next follow-up (10-24 months after first visit)**<br><br>
Table was used to help the project coordinator keeping track of when to remind the clinical sites to have visit 2.<br>
Participants are removed once visit 2 is filled out in REDCap.<br>

```{r, echo=FALSE}

# Display when participants' second visit should occur (start and end dates)
# Only include participants who do not have a second visit reported in REDCap
Visit1[c("bas_patient","centre_name","bas_enrol_mistral_d","vis_stool_samp_rec_d_1")] %>%
  subset(., !(bas_patient %in% Visit2$bas_patient)) %>%
  mutate(Follow_up_1_start = vis_stool_samp_rec_d_1 %m+% months(10)) %>%
  mutate(Follow_up_1_end = vis_stool_samp_rec_d_1 %m+% months(24)) %>%
  datatable(., rownames = FALSE, selection = "multiple",filter = list(position = 'top', clear = TRUE),
            colnames = c("Participant ID","Centre","Enrolment date","Visit 1","Visit 2 start","Visit 2 end"),
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```
**Check time between visit 1 and visit 2**<br><br>
Should be between 10-24 months.<br>

```{r, echo=FALSE}

# Get data from visit 1 for participants who have had their visit 2
V1_d <- Visit1[Visit1$bas_patient %in% Visit2$bas_patient,] %>%
  .[c("bas_patient","centre_name","vis_stool_samp_rec_d_1")] %>%
  rename(Visit1_d = vis_stool_samp_rec_d_1) %>%
  mutate(Visit2_start = Visit1_d %m+% months(10)) %>%
  mutate(Visit2_end = Visit1_d %m+% months(24))

# Get data from visit 2
V2_d <- Visit2 %>%
  .[c("bas_patient","vis_stool_samp_rec_d_1")] %>%
  rename(Visit2_d = vis_stool_samp_rec_d_1) %>%
  subset(., !(is.na(Visit2_d)==T))

# Merge data to get dataframe of visit 1 and 2 dates per participant
V1V2 <- merge(V1_d, V2_d, by="bas_patient")

```

```{r, echo=FALSE}

# Find time difference between visit 1 and 2 per participant
followup_timediff <- V1V2 %>%
  mutate(time_diff = difftime(Visit2_d,Visit1_d))

# Plot timedifference to allow visual examination of outliers
p_follow_up_time <- ggplot(followup_timediff, aes(x = time_diff)) +
  geom_histogram(binwidth = 1, fill=colMISTRAL) +
  xlab("Days between first and second visit") +
  ylab("Count") +
  ggtitle("Time between visits") +
  geom_vline(xintercept=300) +
  geom_text(x=300, y=4, label="10 months", size=3) +
  geom_vline(xintercept=730) +
  geom_text(x=730, y=4, label="24 months", size=3) +
  theme_light(); ggplotly(p_follow_up_time, height=350)

```


<br><br><br>

### Form completeness
**Graphical overview of the completeness of the initiated REDCap forms**<br>

```{r, echo=FALSE}

# Get completion values for each section of visit 1 in REDCap
Visit1_secs <- data.frame("participant"=Visit1$bas_patient,
  "centre"=Visit1$centre_name,
  "baseline"=Visit1$patient_baseline_data_complete.factor,
  "questionnaire"=Visit1$questionnaire_complete.factor, 
  "samples"=Visit1$samples_complete.factor) %>% gather(.,"section","status","baseline":"samples")

# Save plot of completeness
v1secp <- ggplot(Visit1_secs, aes(x=section,fill=status)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  ggtitle("Visit 1") +
  theme(legend.position="none") +
  labs(fill="Status") +
  scale_fill_manual(values=c(Complete="#249225", Incomplete="#cc0000"))

# Get completion values for each section of visit 2 in REDCap
Visit2_secs <- data.frame("participant"=Visit2$bas_patient,
  "centre"=Visit2$centre_name,
  "questionnaire"=Visit2$questionnaire_complete.factor, 
  "samples"=Visit2$samples_complete.factor) %>% gather(.,"section","status","questionnaire","samples")

# Save plot of completeness
v2secp <- ggplot(Visit2_secs, aes(x=section,fill=status)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  ggtitle("Visit 1 and 2") +
  labs(fill="Status") +
  scale_fill_manual(values=c(Complete="#249225", Incomplete="#cc0000"))

# Show two above plots in one figure
plotly::subplot(v1secp, v2secp, nrows = 1, margin = 0.04, shareY = F) %>%
  layout(height = 350); #rm(v1secp, v2secp, v1secp, v2secp)

```

<br><br>

### Gender

```{r Visit1_gender, echo=FALSE}

# Plot reported gender
gender_all <- ggplot(Visit1, aes(x = es_bas_gender.factor, fill = es_bas_gender.factor)) +
  geom_bar() + 
  theme_light() +
  scale_fill_discrete(name = "") +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Gender");ggplotly(gender_all, height=350)

# Plot reported gender per clinical site
gender <- ggplot(Visit1, aes(x=es_bas_gender.factor, fill=es_bas_gender.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=floor(length(unique(Visit1$centre))/4),
             labeller = labeller(centre_name = label_wrap_gen(width=20))) +
  theme_light() +
  scale_fill_discrete(name = "") +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Gender per centre");ggplotly(gender,
                         height=100+(ceiling(length(unique(Visit1$es_bas_gender.factor))/4)*350))

# Remove unnecessary objects
#rm(gender_all, gender)

```

<br><br><br>

### Samples {.tabset}

#### Visit 1

```{r, echo=FALSE}

# Subset those with a completed visit 1 section
Visit1_complete_samples <- Visit1[Visit1$samples_complete == 2,]

# Get variables describing whether a sample was collected
Visit1_samples <- data.frame("participant"=Visit1_complete_samples$bas_patient,
  "centre"=Visit1_complete_samples$centre_name,
  "stool1"=Visit1_complete_samples$es_sample_stool_code_1_h,
  "stool2"=Visit1_complete_samples$es_sample_stool_code_2_h, 
  "stool3"=Visit1_complete_samples$es_sample_stool_code_3_h,
  "stool4"=Visit1_complete_samples$es_sample_stool_code_4_h,
  "stool5"=Visit1_complete_samples$es_sample_stool_code_5_h,
  "stool6"=Visit1_complete_samples$es_sample_stool_code_6_h,
  "plasma1"=Visit1_complete_samples$es_sample_plasma_code_1_h,
  "plasma2"=Visit1_complete_samples$es_sample_plasma_code_2_h, 
  "plasma3"=Visit1_complete_samples$es_sample_plasma_code_3_h,
  "plasma4"=Visit1_complete_samples$es_sample_plasma_code_4_h,
  "plasma5"=Visit1_complete_samples$es_sample_plasma_code_5_h,
  "plasma6"=Visit1_complete_samples$es_sample_plasma_code_6_h,
  "wb1"=Visit1_complete_samples$es_sample_wb_1_h,
  "wb2"=Visit1_complete_samples$es_sample_wb_2_h, 
  "wb3"=Visit1_complete_samples$es_sample_wb_3_h,
  "wb4"=Visit1_complete_samples$es_sample_wb_4_h)

# Find those missing the last stool, plasma or whole blood sample
# (those missing at least 1 sample)
Visit1_missing <- Visit1_samples[Visit1_samples$stool6==0|Visit1_samples$plasma6==0|Visit1_samples$wb4==0,] %>%
  na.omit()

```

**Total participants with status complete:**
`r nrow(Visit1_complete_samples)` <br> 
Out of these, `r nrow(Visit1_missing)` are missing 1 or more aliquots.

Figures show only results from sample sections marked as complete. The graphic was used to track which samples were not collected at visits.<br>

```{r, echo=FALSE}

# Overview of stool aliquots
Visit1_stool <- data.frame("participant"=Visit1_complete_samples$bas_patient,
  "centre"=Visit1_complete_samples$centre_name,
  "stool1"=Visit1_complete_samples$es_sample_stool_code_1_h,
  "stool2"=Visit1_complete_samples$es_sample_stool_code_2_h, 
  "stool3"=Visit1_complete_samples$es_sample_stool_code_3_h,
  "stool4"=Visit1_complete_samples$es_sample_stool_code_4_h,
  "stool5"=Visit1_complete_samples$es_sample_stool_code_5_h,
  "stool6"=Visit1_complete_samples$es_sample_stool_code_6_h) %>%
  gather(.,"sample","status","stool1":"stool6") %>% 
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1stool <- ggplot(Visit1_stool, aes(x=sample,fill=status_yn)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  scale_fill_discrete(name = "") +
  theme(axis.text.x = element_text(angle = 40)) +
  ggtitle("Visit 1: Stool samples") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Overview of plasma aliquots
Visit1_plasma <- data.frame("participant"=Visit1_complete_samples$bas_patient,
  "centre"=Visit1_complete_samples$centre_name,
  "plasma1"=Visit1_complete_samples$es_sample_plasma_code_1_h,
  "plasma2"=Visit1_complete_samples$es_sample_plasma_code_2_h, 
  "plasma3"=Visit1_complete_samples$es_sample_plasma_code_3_h,
  "plasma4"=Visit1_complete_samples$es_sample_plasma_code_4_h,
  "plasma5"=Visit1_complete_samples$es_sample_plasma_code_5_h,
  "plasma6"=Visit1_complete_samples$es_sample_plasma_code_6_h) %>% 
  gather(.,"sample","status","plasma1":"plasma6") %>% 
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1plasma <- ggplot(Visit1_plasma, aes(x=sample,fill=status_yn)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  scale_fill_discrete(name = "") +
  theme(axis.text.x = element_text(angle = 40)) +
  ggtitle("Visit 1: Plasma samples") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Overview of stool samples
Visit1_wb <- data.frame("participant"=Visit1_complete_samples$bas_patient,
  "centre"=Visit1_complete_samples$centre_name,
  "wb1"=Visit1_complete_samples$es_sample_wb_1_h,
  "wb2"=Visit1_complete_samples$es_sample_wb_2_h, 
  "wb3"=Visit1_complete_samples$es_sample_wb_3_h,
  "wb4"=Visit1_complete_samples$es_sample_wb_4_h) %>% 
  gather(.,"sample","status","wb1":"wb4") %>% 
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1wb <- ggplot(Visit1_wb, aes(x=sample,fill=status_yn)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 40)) +
  scale_fill_discrete(name = "") +
  xlab("") +
  ggtitle("Visit 1: Sample aliquots") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Gather plots in one figure
plotly::subplot(v1stool, v1plasma, v1wb, nrows = 1, widths = c(0.38, 0.38, 0.24)) %>%
  layout(height = 350)

```

The heatmap show the participants with 1 or more missing samples in order to see if there were any systematic errors for the clinical sites.<br>

```{r, echo=FALSE}

# Interactive heatmap of those missing at least one sample
heatmaply::heatmaply(Visit1_missing[,c(3:18)],
                     labRow = Visit1_missing$participant,
                     grid_gap = 1,
                     # Set colors
                     colors = c("#cc0000", "#249225"),
                     plot_method= "plotly",
                     # Remove dendrograms
                     Colv=NA,Rowv=NA,
                     row_side_colors = data.frame("Centre"=Visit1_missing$centre),
                     # Remove color bar
                     hide_colorbar = TRUE) %>% layout(height=1500)

```

<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>

NB: Dates were scrambled for the purpose of anonymisation so the time between samples is off compared to the real data.

```{r, echo=FALSE}

ggplotly(ggplot(Visit1, aes(x=floor(difftime(vis_stool_samp_rec_d_1, vis_stool_samp_col_d_1, unit="days")))) +
  geom_bar(stat = "count", fill=colMISTRAL) +
  xlab("Days between stool defacation and recieval at the clinic") +
  ylab("Count") +
  theme_light(), height=400)

```

<br><br><br>

#### Visit 2

```{r, echo=FALSE}

Visit2_complete_samples <- Visit2[Visit2$samples_complete == 2,]

Visit2_samples <- data.frame("participant"=as.factor(Visit2_complete_samples$bas_patient),
                             "centre"=Visit2_complete_samples$centre_name,
                             "stool1"=Visit2_complete_samples$es_sample_stool_code_1_h,
                             "stool2"=Visit2_complete_samples$es_sample_stool_code_2_h,
                             "stool3"=Visit2_complete_samples$es_sample_stool_code_3_h,
                             "stool4"=Visit2_complete_samples$es_sample_stool_code_4_h,
                             "stool5"=Visit2_complete_samples$es_sample_stool_code_5_h,
                             "stool6"=Visit2_complete_samples$es_sample_stool_code_6_h,
                             "plasma1"=Visit2_complete_samples$es_sample_plasma_code_1_h,
                             "plasma2"=Visit2_complete_samples$es_sample_plasma_code_2_h,
                             "plasma3"=Visit2_complete_samples$es_sample_plasma_code_3_h,
                             "plasma4"=Visit2_complete_samples$es_sample_plasma_code_4_h,
                             "plasma5"=Visit2_complete_samples$es_sample_plasma_code_5_h,
                             "plasma6"=Visit2_complete_samples$es_sample_plasma_code_6_h,
                             "wb1"=Visit2_complete_samples$es_sample_wb_1_h,
                             "wb2"=Visit2_complete_samples$es_sample_wb_2_h,
                             "wb3"=Visit2_complete_samples$es_sample_wb_3_h,
                             "wb4"=Visit2_complete_samples$es_sample_wb_4_h)

Visit2_missing <- Visit2_samples[Visit2_samples$stool6==0|Visit2_samples$plasma6==0|Visit2_samples$wb4==0,]

```

**Total participants with status complete:**
`r nrow(Visit2_complete_samples)` <br> Out of these,
`r nrow(Visit2_missing)` are missing 1 or more aliquots.

Figures show only results from sample sections marked as complete.

```{r, echo=FALSE}

#Overview of stool aliquots
Visit2_stool <- data.frame("participant"=Visit2_complete_samples$bas_patient,
                           "centre"=Visit2_complete_samples$centre_name,
                           "stool1"=Visit2_complete_samples$es_sample_stool_code_1_h,
                           "stool2"=Visit2_complete_samples$es_sample_stool_code_2_h,
                           "stool3"=Visit2_complete_samples$es_sample_stool_code_3_h,
                           "stool4"=Visit2_complete_samples$es_sample_stool_code_4_h,
                           "stool5"=Visit2_complete_samples$es_sample_stool_code_5_h,
                           "stool6"=Visit2_complete_samples$es_sample_stool_code_6_h) %>%
  gather(.,"sample","status","stool1":"stool6") %>%
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1stool <- ggplot(Visit2_stool, aes(x=sample,fill=status_yn)) +
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  scale_fill_discrete(name = "") +
  theme(axis.text.x = element_text(angle = 40)) +
  ggtitle("Visit 2: Stool samples") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Overview of plasma aliquots
Visit2_plasma <- data.frame("participant"=Visit2_complete_samples$bas_patient,
  "centre"=Visit2_complete_samples$centre_name,
  "plasma1"=Visit2_complete_samples$es_sample_plasma_code_1_h,
  "plasma2"=Visit2_complete_samples$es_sample_plasma_code_2_h,
  "plasma3"=Visit2_complete_samples$es_sample_plasma_code_3_h,
  "plasma4"=Visit2_complete_samples$es_sample_plasma_code_4_h,
  "plasma5"=Visit2_complete_samples$es_sample_plasma_code_5_h,
  "plasma6"=Visit2_complete_samples$es_sample_plasma_code_6_h) %>%
  gather(.,"sample","status","plasma1":"plasma6") %>%
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1plasma <- ggplot(Visit2_plasma, aes(x=sample,fill=status_yn)) +
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  scale_fill_discrete(name = "") +
  theme(axis.text.x = element_text(angle = 40)) +
  ggtitle("Visit 2: Plasma samples") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Overview of stool samples
Visit2_wb <- data.frame("participant"=Visit2_complete_samples$bas_patient,
  "centre"=Visit2_complete_samples$centre_name,
  "wb1"=Visit2_complete_samples$es_sample_wb_1_h,
  "wb2"=Visit2_complete_samples$es_sample_wb_2_h,
  "wb3"=Visit2_complete_samples$es_sample_wb_3_h,
  "wb4"=Visit2_complete_samples$es_sample_wb_4_h) %>%
  gather(.,"sample","status","wb1":"wb4") %>%
  mutate(status_yn = as.factor(ifelse(status == 11, 'Yes', 'No')))

# Save plot
v1wb <- ggplot(Visit2_wb, aes(x=sample,fill=status_yn)) +
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 40)) +
  scale_fill_discrete(name = "") +
  xlab("") +
  ggtitle("Visit 2: Sample aliquots") +
  scale_fill_manual(values=c(Yes="#249225", No="#cc0000"),name = NULL)

# Gather plot
plotly::subplot(v1stool, v1plasma, v1wb, nrows = 1, widths = c(0.38, 0.38, 0.24)) %>%
  layout(height = 350)

```

The heatmap show the participants with 1 or more missing samples in order to see if there were any systematic errors for the clinical sites.

```{r, echo=FALSE}

# Create interactive heatmap of the missing samples
heatmaply::heatmaply(Visit2_missing[,c(3:18)],
                     labRow = Visit2_missing$participant,
                     grid_gap = 1,
                     # Set colors
                     colors = c("#cc0000", "#249225"),
                     plot_method= "plotly",
                     # Remove dendrograms
                     Colv=NA,Rowv=NA,
                     row_side_colors = data.frame("Centre"=Visit2_missing$centre),
                     # Remove color bar
                     hide_colorbar = TRUE) %>% layout(height=700)

```

<br><br><br><br><br><br><br><br><br><br><br>
NB: Dates were scrambled for the anonymisation so the time between samples is off compared to the real data.

```{r, echo=FALSE}

# Plot days between stool defacation and receival at the clinic
# Should be <=2 days
ggplotly(ggplot(Visit2, aes(x=floor(difftime(vis_stool_samp_rec_d_1, vis_stool_samp_col_d_1, unit="days")))) +
  geom_bar(stat = "count", fill=colMISTRAL) +
  xlab("Days between stool defacation and recieval at the clinic (Visit 2)") +
  theme_light(), height=350)

```

<br><br><br>

### Sample tracking
```{r child='Child_SampleTracking.Rmd'}
# This code block should be empty.
# Another child document is called.
```

<br><br><br>

### Form checks

This section contains tables used to check whether the clinical sites are following the protocol and to do sanity checks to make sure the data entered for the same participants is the same across forms.<br><br>

#### >3 months between consent and samples
Check whether participants have contributed samples and clinical data within a reasonable timeframe.<br>

```{r, echo=FALSE}

# Subset people with stool collection date and display those with >3 months between that and enrolment
# Samples should be collected within 3 months after enrolment
Visit1[!(is.na(Visit1$vis_stool_samp_rec_d_1)),] %>%
  mutate(DateDiff = difftime(vis_stool_samp_rec_d_1, bas_enrol_mistral_d, unit="days")) %>%
  subset(., DateDiff > 90) %>%
  .[c("bas_patient","centre_name","bas_enrol_mistral_d","vis_stool_samp_rec_d_1","DateDiff")] %>%
  datatable(., rownames = FALSE, filter = list(position = 'top', clear = TRUE),
            colnames = c("Participant ID","Centre","Enrolment date","Stool receival date","Time difference (days)"), 
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)
  
```

<br><br>


#### Participants who provided stool before enrolment (consent)
Confirm that all participants follow protocol.

```{r, echo=FALSE}

# Display those who collected stool before enrolment
Visit1[!(is.na(Visit1$vis_stool_samp_rec_d_1)),] %>%
  subset(., as.Date(.$vis_stool_samp_col_d_1) < as.Date(.$bas_enrol_mistral_d)) %>%
  .[c("bas_patient","centre_name","vis_stool_samp_col_d_1","bas_enrol_mistral_d")] %>%
  datatable(., rownames = FALSE, filter = list(position = 'top', clear = TRUE),
            colnames = c("Participant ID","Centre","Stool defacation date","Enrolment date"),
            options = list(dom = 't', columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```

<br><br>

#### Birthdates that do not match with **enrolment form**
Check that entered birthdates match between different REDCap forms.

```{r, echo=FALSE}

# Subset those with both a clinical and a visit form filled out
Clinical_subset <- Clinical[(Clinical$bas_patient %in% Visit1$bas_patient),]
Visit1_subset <- Visit1[(Visit1$bas_patient %in% Clinical_subset$bas_patient),]

# Subset birthdates from each REDCap form
Birthd_subset <- data.frame("bas_patient"    = Clinical_subset$bas_patient,
                            "clinical_bd"    = Clinical_subset$bas_birth_d,
                            "visit_bd"       = Visit1_subset$es_bas_birth_d) %>%
  mutate(Error = ifelse(clinical_bd==visit_bd, 0,1))

# Display those where birthdates do not match across forms
subset(Birthd_subset,Error==1) %>%
  select(.,-c("Error")) %>%
  datatable(., rownames = F, filter = list(position = 'top', clear = TRUE), class = table_style,
            colnames = c("Participant ID","Clinical b-day","Visit1 b-day"),
            options = list(dom = 't', columnDefs = list(list(className = 'dt-center', targets = "_all"))))

```

<br><br>

#### Gender that does not match with **Clinical form**
Check that entered birthdates match between different REDCap forms.<br>
*Table is empty because no such errors were found in the data.*

```{r, echo=FALSE}

# Using dataframes of participants with both clinical and visit forms filled out
# subset gender in both dataframes
Gender_subset <- data.frame(Clinical_subset$bas_patient,
                            Clinical_subset$es_bas_gender.factor,
                            Visit1_subset$es_bas_gender.factor) %>%
  mutate(Error = ifelse(Clinical_subset$es_bas_gender.factor==Visit1_subset$es_bas_gender.factor,0,1))

# Display those with a non-matching gender across forms
subset(Gender_subset, Error==1) %>%
  select(.,-c("Error")) %>%
  datatable(., rownames = F, filter = list(position = 'top', clear = TRUE),
            colnames = c("Participant ID","Enrolment gender","Visit1 gender"),
            options = list(dom = 't', columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```


<br><br><br>


### Alcohol and exercise

Sanity checks of selected variables.<br>
*The "Alcohol consumption reported in Visit 1" looks weird because of an outlier which was detected because of this graphic.*

```{r, echo=FALSE}

# Plot exercise per week reported at visit 1
p_exer1 <- ggplot(Visit1, aes(x = vis_exercise_1m_fr)) + 
  geom_histogram(binwidth = 1, fill=colMISTRAL) +
  xlab("Average exercise (hours/week)") +
  ylab("Count") +
  ggtitle("Exercise reported in Visit 1") +
  scale_x_continuous(breaks = seq(0, max(Visit1$vis_exercise_1m_fr, na.rm = TRUE), 5)) +
  theme_light(); ggplotly(p_exer1, height=350)

# Plot exercise per week reported at visit 2
p_exer2 <- ggplot(Visit2, aes(x = vis_exercise_1m_fr)) + 
  geom_histogram(binwidth = 1, fill=colMISTRAL) +
  xlab("Average exercise (hours/week)") +
  ylab("Count") +
  ggtitle("Exercise reported in Visit 1 and 2") +
  scale_x_continuous(breaks = seq(0, max(Visit2$vis_exercise_1m_fr, na.rm = TRUE), 5)) +
  theme_light(); ggplotly(p_exer2, height=350)

# Plot alcohol units per week reported at visit 1
p_alco1 <- ggplot(Visit1, aes(x = vis_subs_alco_1m_fr)) + 
  geom_histogram(binwidth = 1, fill=colMISTRAL) +
  xlab("Average alcohol consumption (units/week)") +
  ylab("Count") +
  ggtitle("Alcohol consumption reported in Visit 1") +
  scale_x_continuous(breaks = seq(0, max(Visit1$vis_subs_alco_1m_fr, na.rm = TRUE), 2)) +
  theme_light(); ggplotly(p_alco1, height=350)

# Plot alcohol units per week reported at visit 2
p_alco2 <- ggplot(Visit2, aes(x = vis_subs_alco_1m_fr)) + 
  geom_histogram(binwidth = 1, fill=colMISTRAL) +
  xlab("Average alcohol consumption (units/week)") +
  ylab("Count") +
  ggtitle("Alcohol consumption reported in Visit 2") +
  scale_x_continuous(breaks = seq(0, max(Visit2$vis_subs_alco_1m_fr, na.rm = TRUE), 2)) +
  theme_light(); ggplotly(p_alco2, height=350)

```

<br><br><br>

### Genome analysis
The participants were asked for consent to genomics analysis.

#### Did the participant consent to host genomics analysis
The empty row represents unanswered questions.

```{r, echo=FALSE}

# Count participant consents and calculate percentage relative to all enrolled participants
# Display counts and percentages
Visit1 %>% count(mist_bas_genom_cons.factor) %>% mutate(percent = round(n/sum(n)*100,1)) %>%
  datatable(., rownames=FALSE, colnames=c("Genomics analysis?","n","%"), class = table_style, 
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))))

# Plot consent results per clinical site
p_genom <- Visit1 %>% count(mist_bas_genom_cons.factor,centre) %>% mutate(percent = round(n/sum(n)*100,1)) %>%
  ggplot(., aes(x = mist_bas_genom_cons.factor, y = n, fill=mist_bas_genom_cons.factor)) +
  geom_bar(stat = "identity") +
  facet_wrap("centre") + 
  xlab("Consent response") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Consent to genomics analysis per centre") +
  scale_fill_manual("",values=c("#249225", "#cc0000", "#dddddd")) +
  theme_light(); ggplotly(p_genom, height = 600, width = 800)

```

<br><br><br>

### Diet
Assembled grapic of data collected in different variables (omnivore yes/no, pescetarian yes/no etc.).

```{r fun facts, echo=FALSE, fig.width = 6, fig.asp = .6}

# Diet variable created from multiple options
Visit1_diet <- Visit1 %>% mutate(Diet = case_when(vis_diet_t_omni___omni == 1 ~ "Omnivore",
                                                  vis_diet_t_pesc___pesc == 1 ~ "Pescetarian",
                                                  vis_diet_t_vege___vege == 1 ~ "Vegetarian",
                                                  vis_diet_t_vega___vega == 1 ~ "Vegan"))

# Plot diet 
ggplotly(ggplot(Visit1_diet, aes(Diet, fill=Diet)) +
  geom_bar(stat="count") +
  scale_fill_discrete(name = "") +
  ggtitle("Diet") +
  xlab("Diet classification") +
  ylab("Count") +
  theme_light(), height=350)

```

<br><br><br>

```{r, include=FALSE}
# Clear memory
gc()
```

