---
title: "Child_Clinical"
author: "Kirstine K. Rasmussen & Sophia Hejndorf"
date: "`r Sys.Date()`"
output: html_document
---
## **Clinical form** {.tabset}

*The Clinical form was used for collection patient clinical and historical health information. Data collected here includes patient demographic information, HIV-related clinical data, recent laboratory values, immune cell measurements, antiretroviral treatment information and clinical events. This form did not require the participant to come in for a visit.*

**Total participants found in the Clinical form:** `r nrow(Clinical)`

### General overview

**Completion date for section** <br>
Min: `r min(Clinical$es_completed_date_a2)` <br>
Median: `r median(Clinical$es_completed_date_a2)` <br>
Max: `r max(Clinical$es_completed_date_a2)`

**Age of participants at enrolment** <br>
Min: `r summary(as.integer(Clinical$age_at_enrolment))[1]` <br>
Median: `r summary(as.integer(Clinical$age_at_enrolment))[3]` <br>
Max: `r summary(as.integer(Clinical$age_at_enrolment))[6]`

```{r, echo=FALSE}

# Plot participants with Clinical forms per site
enroledp <- ggplot(Clinical, aes(x=centre,fill=centre_name)) + 
  geom_bar() +
  xlab("Centre") +
  ylab("Count") +
  theme_light() +
  labs(fill="Centre name") +
  ggtitle("Participants with Clinical forms per site");ggplotly(enroledp, height=350)

```

<br><br><br>

### Form completeness

```{r, echo=FALSE}

# Completeness of Clinical forms

# Get completion status per section in the form
sections <- data.frame("participant"=Clinical$bas_patient,
  "centre"=Clinical$centre_name,
  "a1 (demograpgy)"=Clinical$section_a1_demography_and_hiv_status_complete.factor,
  "a2 (clinical)"=Clinical$section_a2_basic_clinical_information_complete.factor, 
  "b1 (lab)"=Clinical$section_b1_laboratory_values_complete.factor, 
  "b2 (CD4, CD8, HIV)"=Clinical$section_b2_cd4_cd8_and_hivrna_complete.factor,
  "b3 (HCV)"=Clinical$section_b3_hepatitis_virology_and_fibrosis_screeni_complete.factor,
  "b4 (COVID"=Clinical$section_b4_covid19_complete.factor,
  "c1 (ART)"=Clinical$section_c1_antiretroviral_treatment_complete.factor,
  "c2 (CVD)"=Clinical$section_c2_medication_related_to_cardiovascular_di_complete.factor,
  "c3 (HCV treatment)"=Clinical$section_c3_treatment_against_hepatitis_c_complete.factor,
  "d (opportunistic inf.)"=Clinical$section_d_severe_opportunistic_infections_and_sexu_complete.factor) %>% gather(.,"section","status","a1..demograpgy.":"d..opportunistic.inf..")

# Plot completions
secp <- ggplot(sections, aes(x=section,fill=status)) + 
  geom_bar(position = position_stack(reverse=T)) +
  theme_light() +
  xlab("Section") +
  ylab("Count") +
  labs(fill="Status") +
  theme(axis.text.x = element_text(angle = 300)) +
  ggtitle("Clinical: Form completeness by section") +
  scale_fill_manual(values=c(Complete=colYes, Incomplete=colNo, Unverified="#dddddd"));ggplotly(secp, height=400)

```

#### Clinical form status

```{r, echo = FALSE}

# Display overall completion status for Clinical form
Clinical %>% 
  group_by(status_complete.factor) %>% 
  summarise(n=n()) %>% 
  adorn_totals(where = "row") %>% 
  datatable(., rownames = F, colnames = c("Status","n"), caption = "Clinical form status",
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

# Display overall completion status for Clinical form per clinical site
Clinical  %>%
  group_by(centre, centre_name, status_complete.factor) %>% 
  summarise(n=n()) %>% 
  adorn_totals(where = "row") %>% 
  datatable(., rownames = F, colnames = c("Centre","Centre name","Status","n"), caption = "Clinical form status per centre",
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```

<br><br><br>

### Other ART medications

ART treatments listed under "other" instead of selected in our list of ART drugs. Each ART regiment was written in their own field from a drop-down or by selecting "other". This is why many fields are empty - they were filled by a drug on the defined list and were not an issue.

```{r, echo=FALSE}

# Get all ART treatments listed as "other" (not in the defined list)
ART <- Clinical[c("bas_patient","centre",
                  "es_art_id_oth_1","es_art_id_oth_2","es_art_id_oth_3","es_art_id_oth_4")]
# Show only the ones with at least one ART treatment listed under "other"
ART[rowSums(is.na(ART)) != ncol(ART)-2, ] %>% 
  datatable(., rownames = FALSE, class = table_style, 
            colnames = c("Participant ID","Centre","Other drug 1","Other drug 2","Other drug 3","Other drug 4"),
            filter = list(position = 'top', clear = TRUE),
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))))

```

<br><br><br>

### Missing data and errors

Only forms marked as complete are included below.

#### Participants with no height / weight in the Clinical form

```{r complete_forms, echo=FALSE}

# Subset completed forms
complete <- subset(Clinical, status_complete.factor=="Complete")
# Subset those where height or weight is missing (NA) or reported as missing (999)
# Display missingness
subset(complete, is.na(es_bas_weigh) | es_bas_weigh==999 | is.na(bas_heigh) | bas_heigh==999) %>% 
  .[c("bas_patient","centre","bas_heigh","es_bas_weigh","centre_name")] %>%
  datatable(.,rownames=FALSE, class = table_style, 
            colnames = c("Participant ID","Centre","Height","Weight","Centre name"),
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))))

```

<br><br><br>

#### Participants who tested positive for HIV-1 after their first visit to the clinic
Check for discrepancies in the entered data.

```{r, echo=FALSE}

# Display those where the first positive HIV test date is before the first reported visit to the clinic
subset(complete,complete$es_lab_viro_vs_d_hiv_1_pos>complete$bas_frsvis_d) %>% .[c("bas_patient","es_lab_viro_vs_d_hiv_1_pos","bas_frsvis_d","centre_name")] %>%
  datatable(., rownames = FALSE, colnames = c("Participant ID","First positive HIV-1","First visit to clinic","Site"), 
            filter = list(position = 'top', clear = TRUE),
            class = table_style, options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))))

```
<br><br><br>

### Lab values {.tabset}
Important lab values were explored with regard to how many had entered values for the measurement, the distribution of the measurements, and whether any measurements were outside the expected (different from normal) range.

```{r Custum functions, echo=FALSE}

# Function for generating lab value plots
plot_generator <- function(data, unit, lower_limit, upper_limit, xmin, xmax, bin, caption){
 
  p <- ggplotly(ggplot(subset(data, LAB_U_ == unit), aes(x = LAB_V_)) + 
  geom_histogram(binwidth = bin, fill = colMISTRAL) +
  ylab("Count") +
  labs(title = caption) +
  geom_vline(xintercept = lower_limit, color = colNA) +
  geom_vline(xintercept = upper_limit, color = colNA) +
  scale_x_continuous(limits = c(xmin, xmax)) +
  theme_light(), height = 350)
  
  return(p)
}

# Function for generating a table of how many lab values are below, within or above the limits
table_generator <- function(data, unit, lower_limit, upper_limit, caption){

  data2 <- data %>% drop_na(LAB_V_)
  
  Category <- c("Below range", "Within range", "Above range")
  
  Count <- c(sum(data2$LAB_U_ == unit & data2$LAB_V_ < lower_limit),
             sum(data2$LAB_U_ == unit & data2$LAB_V_ > lower_limit & data2$LAB_V_ < upper_limit),
             sum(data2$LAB_U_ == unit & data2$LAB_V_ > upper_limit))
  
  Fraction <- c(mean(data2$LAB_V_[data2$LAB_U_ == unit] < lower_limit),
                mean(data2$LAB_V_[data2$LAB_U_ == unit] > lower_limit & data2$LAB_V_[data2$LAB_U_ == unit] < upper_limit),
                mean(data2$LAB_V_[data2$LAB_U_ == unit] > upper_limit))
  
  df <- datatable(data.frame(Category, Count, Fraction),
                  options = list(dom = 't'), class = table_style,
                  caption = caption)
  
  return(df)
}

```

#### ALT

```{r, echo=FALSE}

# Create ALT dataframe
alt <- Clinical %>%
 select(matches("bas_patient|lab_alt|_d_all")) %>%
 select(!(matches(".factor|_2|_3")))

# Subset data from collected samples (1=yes)
alt1 <- alt %>%
 subset(es_lab_alt_y == 1)

# Make date column based on defined date and if none, the date for all lab tests
alt1_d <- alt1 %>%
  mutate(LAB_D_ = ifelse(is.na(lab_alt_d_1)==F, lab_alt_d_1, es_labtests_d_all)) %>%
  mutate(LAB_D_ = as.Date(LAB_D_, origin = "1970-01-01"))

# Select columns, add unit column, change column names
alt1_d_u_v <- alt1_d %>% select(bas_patient, LAB_D_, lab_alt_v_ul_1) %>%
                          mutate(LAB_U_ = "IU/L") %>%
                          rename(c("LAB_V_" = "lab_alt_v_ul_1"));rm(alt,alt1,alt1_d)
  
# Create plot and table from functions
alt_p1 <- plot_generator(data = alt1_d_u_v, unit = "IU/L", lower_limit = 5, upper_limit = 5000, xmin = 0, xmax = 100, bin = 2, caption = "ALT, IU/L")
alt_t1 <- table_generator(data = alt1_d_u_v, unit = "IU/L", lower_limit = 5, upper_limit = 5000, caption = "ALT, IU/L")

```

```{r, echo=FALSE}

# Display how many times the lab value has been measured
datatable(data.frame(table(Clinical$es_lab_alt_y.factor, useNA = "always")),
                  colnames=c("Total response to 'Has ALT (Alanine Aminotransferase) been measured?'", "Count"),
                  options = list(dom = 't'),class = table_style,
                  rownames = FALSE)

# Plot how many times the lab value has been measured per clinical site
alt_y_p <- ggplot(Clinical, aes(x=es_lab_alt_y.factor, fill=es_lab_alt_y.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=ceiling(length(unique(Clinical$es_lab_alt_y.factor))/4)) +
  theme_light() +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values=c(Yes = colYes, No = colNo, Unknown= colUnknown,
                            "NA" = colNA), na.value = colNA,name = NULL) +
  ggtitle("Response to 'Has ALT been measured?' per site");ggplotly(alt_y_p,
                          height=100+(ceiling(length(unique(Clinical$es_lab_alt_y.factor))/4)*250))

# Plot distribution
ggplotly(alt_p1) %>%
  layout(title = "Distribution of all ALT values",xaxis = list(title = "IU/L"))
# Display table
alt_t1

# Remove objects no longer needed
rm(alt, alt_p1, alt_t1, alt_y_p, alt1_d, alt1_d_u_v, alt_long, alt_1_u_d)

```

<br><br>

#### Platelets
```{r, echo=FALSE}

# Create platelets dataframe
thr <- Clinical %>%
 select(matches("bas_patient|lab_thr|_d_all")) %>%
 select(!(matches(".factor|_2|_3")))

# Subset data from collected samples (1=yes)
thr1 <- thr %>%
 subset(es_lab_thr_y == 1)

# Make date column based on defined date and if none, the date for all lab tests
thr1_d <- thr1 %>%
  mutate(LAB_D_ = ifelse(is.na(es_lab_thr_d_1)==F, es_lab_thr_d_1, es_labtests_d_all)) %>%
  mutate(LAB_D_ = as.Date(LAB_D_, origin = "1970-01-01"))

# select columns, add unit column, change column names
thr1_d_u_v <- thr1_d %>% select(bas_patient, LAB_D_, es_lab_thr_v_1) %>%
                          mutate(LAB_U_ = "10^9/L") %>%
                          rename(c("LAB_V_" = "es_lab_thr_v_1")); rm(thr, thr1, thr1_d)
  
# create plot and table
thr_p1 <- plot_generator(data = thr1_d_u_v, unit = "10^9/L", lower_limit = 5, upper_limit = 1000, xmin = -10, xmax = 1010, bin = 10, caption = "Platelets, IU/L")
thr_t1 <- table_generator(data = thr1_d_u_v, unit = "10^9/L", lower_limit = 5, upper_limit = 1000, caption = "Platelets, IU/L")
```

```{r, echo=FALSE}

datatable(data.frame(table(Clinical$es_lab_thr_y.factor, useNA = "always")),
                  colnames=c("Total response to 'Has the platelet count been measured?'", "Count"),
                  options = list(dom = 't'),class = table_style,
                  rownames = FALSE)

thr_y_p <- ggplot(Clinical, aes(x=es_lab_thr_y.factor, fill=es_lab_thr_y.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=ceiling(length(unique(Clinical$es_lab_thr_y.factor))/4)) +
  theme_light() +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values=c(Yes = colYes, No = colNo, Unknown= colUnknown,
                            "NA" = colNA), na.value = colNA,name = NULL) +
  ggtitle("Response to 'Has the platelet count been measured?' per site");ggplotly(thr_y_p,
                         height=100+(ceiling(length(unique(Clinical$es_lab_thr_y.factor))/4)*250))
 
 
# print distribution plot and table
ggplotly(thr_p1) %>%
  layout(title = "Distribution of all platelet counts",xaxis = list(title = "cells/nL"))
thr_t1

```

<br><br>

#### Cholesterol

```{r, echo=FALSE}
# Create chol dataframe
chol <- Clinical %>%
 select(matches("bas_patient|chol|_d_all")) %>%
 select(!(matches(".factor|_2|_3")))

# Subset data from collected samples (1=yes)
chol1 <- chol %>%
 subset(lab_chol_y == 1)

# Make date column based on defined date and if none, the date for all lab tests
chol1_d <- chol1 %>%
  mutate(LAB_D_ = ifelse(is.na(lab_chol_d_1)==F, lab_chol_d_1, es_labtests_d_all)) %>%
  mutate(LAB_D_ = as.Date(LAB_D_, origin = "1970-01-01"))

chol1_d_v_u <- chol1_d %>%
  mutate(LAB_V_ = ifelse(is.na(lab_chol_v_mmol_1)==F, lab_chol_v_mmol_1, lab_chol_v_mgdl_1)) %>%
  mutate(LAB_U_ = lab_chol_u_1); rm(chol, chol1, chol1_d)

chol_p1 <- plot_generator(data = chol1_d_v_u, unit = "mmol_L", lower_limit = 1.8, upper_limit = 11, xmin = 0, xmax = 12, bin = 0.5, caption = "Cholesterol, mmol/L")
chol_t1 <- table_generator(data = chol1_d_v_u, unit = "mmol_L", lower_limit = 1.8, upper_limit = 11, caption = "Cholesterol, mmol/L")

chol_p2 <- plot_generator(data = chol1_d_v_u, unit = "mg_dL", lower_limit = 40, upper_limit = 400, xmin = 0, xmax = 430, bin = 10, caption = "Cholesterol, mg/dL")
chol_t2 <- table_generator(data = chol1_d_v_u, unit = "mg_dL", lower_limit = 40, upper_limit = 400, caption = "Cholesterol, mg/dL")

```

```{r, echo=FALSE}

datatable(data.frame(table(Clinical$lab_chol_y.factor, useNA = "always")),
                  colnames=c("Response to 'Has the total cholesterol been measured?'", "Count"),
                  options = list(dom = 't'),class = table_style,
                  rownames = FALSE)

chol_y_p <- ggplot(Clinical, aes(x=lab_chol_y.factor, fill=lab_chol_y.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=ceiling(length(unique(Clinical$lab_chol_y.factor))/4)) +
  theme_light() +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values=c(Yes = colYes, No = colNo, Unknown= colUnknown,
                             "NA" = colNA), na.value = colNA,name = NULL) +
  ggtitle("Response to 'Has the total cholesterol been measured?' per site");ggplotly(chol_y_p,
                         height=100+(ceiling(length(unique(Clinical$lab_chol_y.factor))/4)*250))

# print plot and table
plotly::subplot(chol_p1, chol_p2) %>%
  layout(title = "Distribution of all cholesterol values",
         xaxis = list(title = "mmol/L"), xaxis2 = list(title = "mg/dL"),
         yaxis = list(title = "Count"), yaxis2 = list(title = ""))

chol_t1
chol_t2

```

<br><br>

#### Haemoglobin

```{r, echo=FALSE}

datatable(data.frame(table(Clinical$es_lab_haem_y.factor, useNA = "always")),
                  colnames=c("Total response to 'Has haemoglobin been measured?'", "Count"),
                  options = list(dom = 't'),class = table_style,
                  rownames = FALSE)

ggplotly(ggplot(Clinical, aes(x=es_lab_haem_y.factor, fill=es_lab_haem_y.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=1) +
  theme_light() +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values=c(Yes = colYes, No = colNo, Unknown= colUnknown,
                             "NA" = colNA), na.value = colNA,name = NULL) +
  ggtitle("Response to 'Has haemoglobin been measured?' per site"),height=350)

```


```{r, echo=FALSE}

haem_p1 <- ggplotly(ggplot(subset(Clinical, lab_haem_u_1.factor == "mmol/L"), aes(x = lab_haem_v_mmol_1)) + 
  geom_histogram(binwidth = 0.25, fill = colMISTRAL) +#, , alpha = 0.5 fill = "darkgrey") +
  labs(title = "Haemoglobin, mmol/L") +
  ylab("Count") +
  geom_vline(xintercept = 6.8, color = colNA) +
  geom_vline(xintercept = 11, color = colNA) +
  theme_light(), height = 350)

haem_p2 <- ggplotly(ggplot(subset(Clinical, lab_haem_u_1.factor == "g/L"), aes(x = lab_haem_v_gl_1)) + 
  geom_histogram(binwidth = 10, fill = colMISTRAL) +#, , alpha = 0.5 fill = "darkgrey") +
  labs(title = "Haemoglobin, g/L") +
  ylab("Count") +
  geom_vline(xintercept = 110, color = colNA) +
  geom_vline(xintercept = 174, color = colNA) +
  theme_light(), height = 350)

plotly::subplot(haem_p1, haem_p2) %>%
  layout(title = "Distribution of all haemoglobin values", xaxis = list(title = "mmol/L"), xaxis2 = list(title = "g/L"))

```


```{r, echo=FALSE}

subset(Clinical, lab_haem_u_1.factor == "mmol/L") %>% 
  .[.$lab_haem_v_mmol_1<6.8 | .$lab_haem_v_mmol_1>11 ,] %>%
  select(c("centre_name","bas_patient","lab_haem_v_mmol_1")) %>%
  datatable(., colnames=c("Centre","PID"," Haemoglobin (mmol/L)"), 
            class = table_style, rownames = FALSE)

subset(Clinical, lab_haem_u_1.factor == "g/L") %>% 
  .[.$lab_haem_v_gl_1<110 | .$lab_haem_v_gl_1>174 ,] %>%
  select(c("centre_name","bas_patient","lab_haem_v_gl_1")) %>%
  datatable(., colnames=c("Centre","PID","Haemoglobin (g/L)"), 
            class = table_style, rownames = FALSE)

```

<br><br>

#### Hba1C

```{r, echo=FALSE}

# Create hba1c dataframe
hba1c <- Clinical %>%
  select(matches("bas_patient|es_lab_hba1c|_d_all")) %>%
  select(!(matches(".factor")))

# Subset data from collected samples (1=yes)
hba1c1 <- hba1c %>%
 subset(es_lab_hba1c_y == 1)

# Make date column based on defined date and if none, the date for all lab tests
hba1c1_d <- hba1c1 %>%
  mutate(LAB_D_ = ifelse(is.na(es_lab_hba1c_d_1)==F, es_lab_hba1c_d_1, es_labtests_d_all)) %>%
  mutate(LAB_D_ = as.Date(LAB_D_, origin = "1970-01-01"))

hba1c1_d_u_v <- hba1c1_d %>%
  mutate(LAB_V_ = ifelse(is.na(es_lab_hba1c_v_mmolm)==F, es_lab_hba1c_v_mmolm, es_lab_hba1c_v_per)) %>%
  mutate(LAB_U_ = es_lab_hba1c_u_1);rm(hba1c,hba1c1,hba1c1_d)

h_p1 <- plot_generator(data = hba1c1_d_u_v, unit = "mmolm", lower_limit = 20, upper_limit = 110, xmin = 5, xmax = 120, bin = 5, caption = "HbA1C, mmol/L")
h_t1 <- table_generator(data = hba1c1_d_u_v, unit = "mmolm", lower_limit = 20, upper_limit = 110, caption = "HbA1C, mmol/L")

h_p2 <- plot_generator(data = hba1c1_d_u_v, unit = "percentage", lower_limit = 4, upper_limit = 12, xmin = 2, xmax = 15, bin = 0.5, caption = "HbA1C, %")
h_t2 <- table_generator(data = hba1c1_d_u_v, unit = "percentage", lower_limit = 4, upper_limit = 12, caption = "HbA1C, %")
 
```

```{r, echo=FALSE}

# Assess answers for sample collection
datatable(data.frame(table(Clinical$es_lab_hba1c_y.factor, useNA = "always")),
          colnames=c("Total response to 'Has HbA1C been measured?'", "Count"),
          options = list(dom = 't'),class = table_style,
          rownames = FALSE)

hba1c_y_p <- ggplot(Clinical, aes(x=es_lab_hba1c_y.factor, fill=es_lab_hba1c_y.factor)) +
  geom_bar() +
  facet_wrap(~ centre, nrow=ceiling(length(unique(Clinical$es_lab_hba1c_y.factor))/4)) +
  theme_light() +
  xlab("") +
  ylab("Count") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values=c(Yes = colYes, No = colNo, Unknown= colUnknown,
                             "NA" = colNA), na.value = colNA,name = NULL) +
  ggtitle("Response to 'Has HbA1C been measured?' per site");ggplotly(hba1c_y_p,
                         height=100+(ceiling(length(unique(Clinical$es_lab_hba1c_y.factor))/4)*250))

plotly::subplot(h_p1, h_p2) %>%
  layout(title = "Distribution of all HbA1C values", xaxis = list(title = "mmol/mol"), xaxis2 = list(title = "Percentage"))

h_t1
h_t2

```

<br><br><br>

### CD4

```{r, echo=FALSE}

# Select all values relating to CD4 cell counts
CD4 <- Clinical %>%
 select(matches("bas_patient|_cd4_")) %>%
 select(!(matches(".factor")))

# Gather units from unit "all column" if not specified
CD4_u <- CD4 %>%
 mutate(across(grep("_u_[1-9]", names(.)),
               ~ ifelse(is.na(.)==FALSE, ., es_lab_cd4_u_all)))

# Gather values from all value columns
CD4_u_v <- CD4_u %>%
  mutate(es_lab_cd4_le_v = ifelse(is.na(es_lab_cd4_le_v_cellmc_1) == F, 
                                  es_lab_cd4_le_v_cellmc_1, es_lab_cd4_le_v_109l_1)) %>%
  mutate(es_lab_cd4_artstart_v = ifelse(is.na(es_lab_cd4_artstart_v_cellmc_1) == F,
                                        es_lab_cd4_artstart_v_cellmc_1, es_lab_cd4_artstart_v_109l_1))

# Create plot of nadir CD4 cell counts in one unit
CD4_nadir_p1 <- ggplotly(ggplot(subset(CD4_u_v, es_lab_cd4_le_u_1 == "10_9_L"), aes(x = es_lab_cd4_le_v)) + 
  geom_histogram(binwidth = 0.1, fill = colCD4) +
  ylab("Count") +
  theme_light(), height = 350)

# Create plot of nadir CD4 cell counts in other reported unit
CD4_nadir_p2 <- ggplotly(ggplot(subset(CD4_u_v, es_lab_cd4_le_u_1 == "cells_mcl"), aes(x = es_lab_cd4_le_v)) + 
  geom_histogram(binwidth = 50, fill = colCD4) +
  ylab("Count") +
  theme_light(), height = 350)

# Create plot of CD4 cell counts at ART initiation in one unit
CD4_ART_p1 <- ggplotly(ggplot(subset(CD4_u_v, es_lab_cd4_artstart_u_1 == "10_9_L"), aes(x = es_lab_cd4_artstart_v)) + 
  geom_histogram(binwidth = 0.1, fill = colCD4) +
  ylab("Count") +
  theme_light(), height = 350)

# Create plot of CD4 cell counts at ART initiation in other reported unit
CD4_ART_p2 <- ggplotly(ggplot(subset(CD4_u_v, es_lab_cd4_artstart_u_1 == "cells_mcl"), aes(x = es_lab_cd4_artstart_v)) + 
  geom_histogram(binwidth = 50, fill = colCD4) +
  ylab("Count") +
  theme_light(), height = 350)

# Plot nadir CD4 cell count plots in one figure
plotly::subplot(CD4_nadir_p1, CD4_nadir_p2) %>%
  layout(title = "Nadir CD4 cell counts", margin = list(t = 50),
    xaxis = list(title = "cells/nL"), xaxis2 = list(title = "cells/mcl"))

# Plot CD4 cell count at ART initiation plots in one figure
plotly::subplot(CD4_ART_p1, CD4_ART_p2) %>%
  layout(title = "CD4 cell counts at ART initiation", margin = list(t = 50),
    xaxis = list(title = "cells/nL"), xaxis2 = list(title = "cells/mcl"))

# Display those with CD4 cell counts outside limits
subset(CD4_u_v, es_lab_cd4_artstart_u_1 == "10_9_L" & !(is.na(es_lab_cd4_artstart_v_109l_1))) %>% .[.$es_lab_cd4_artstart_v > 10,] %>%
.[c("bas_patient","es_lab_cd4_artstart_d_1","es_lab_cd4_artstart_u_1","es_lab_cd4_artstart_v_109l_1")] %>%
  datatable(., rownames = FALSE, caption = "Outliers for CD4 cell counts at ART initiation (cells/nL)", 
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```

```{r, echo=F}

# Display those with reported unknown CD4 cell counts at ART initiation
Clinical[c("bas_patient","centre","es_lab_cd4_le_d_1","es_lab_cd4_artstart_d_1")] %>%
  filter_all(any_vars(. %like% c("1911"))) %>%
  datatable(., rownames = FALSE, colnames = c("PID","Site","Nadir CD4 date","ART init CD4 date"), 
            caption = "Unknown dates for CD4 at ART init", 
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

# Display those with not filled out CD4 cell counts at ART initiation
Clinical[c("bas_patient","centre","es_lab_cd4_le_d_1","es_lab_cd4_artstart_d_1")] %>%
  filter_all(any_vars(is.na(.))) %>%
  datatable(., rownames = FALSE, colnames = c("PID","Site","Nadir CD4 date","ART init CD4 date"), 
            caption = "Blank dates for CD4 at ART init", 
            options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))), class = table_style)

```

<br>